Modeling HEPES buffer with a charge-balance approach

### 2023-12-01 noise.Rmd

Every row of the newly obtained `data.frame` now contains results for
every repeat of the same set of experimental conditions:

@@

This lets us plot the average estimation error and the standard
deviation of the estimates:
```{r}
lattice::levelplot(
	res_mean ~ noise_ratio + n_points | noise_kind, pna,
	group = noise_kind, scales = list(log = 10),
	main = quote(
		'Average '*pK[a]*' estimation error '*
		'(should be close to 0)'
	),
	xlab = 'Noise-to-signal ratio',
	ylab = 'Number of known points',
	cuts = 255, col.regions = hcl.colors(256, 'Broc')
)
```

I don't see any pattern in the data, and it seems to be fairly centered
around $0$, so the estimate is effectively unbiased, even in presence of
non-normally-distributed noise. But what about the variance?
```{r}
lattice::levelplot(
	sqrt(res_dev) ~ noise_ratio + n_points | noise_kind, pna,
	group = noise_kind, scales = list(log = 10),
	main = quote(
		'Standard deviation of '*pK[a]*' estimate'
	),
	xlab = 'Noise-to-signal ratio',
	ylab = 'Number of known points',
	cuts = 255, col.regions = hcl.colors(256, 'Terrain 2')
)
```

Given $10^{1.5}$ points and realistic (?) noise, the standard deviation
of the $\mathrm{p}K_a$ estimates can be as high as $1.5$. Under
three-sigma rule, this gives us an error of up to $4.5$ $\mathrm{pH}$
units...

>>>>

Every row of the newly obtained `data.frame` now contains results for
all repeats of the same set of experimental conditions:

@@

This lets us plot the average estimation error and the standard
deviation of the estimates:
```{r}
library(lattice)
levelplot(
	res_mean ~ noise_ratio + n_points | noise_kind, pna,
	group = noise_kind, scales = list(log = 10),
	main = quote(
		'Average '*pK[a]*' estimation error '*
		'(should be close to 0)'
	),
	xlab = 'Noise-to-signal ratio',
	ylab = 'Number of known points',
	cuts = 255, col.regions = hcl.colors(256, 'Broc')
)
```

I don't see any pattern in the data, and it seems to be fairly centered
around $0$, so the estimate is effectively unbiased, even in presence of
non-normally-distributed noise. But what about the variance?
```{r}
levelplot(
	sqrt(res_dev) ~ noise_ratio + n_points | noise_kind, pna,
	group = noise_kind, scales = list(log = 10),
	main = quote(
		'Standard deviation of '*pK[a]*' estimate'
	),
	xlab = 'Noise-to-signal ratio',
	ylab = 'Number of known points',
	cuts = 255, col.regions = hcl.colors(256, 'Terrain 2'),
	panel = function(x, y, z, subscripts, at, contour, region, ...) {
		panel.levelplot(
			x = x, y = y, z = z, subscripts = subscripts,
			at = at, contour = FALSE, region = TRUE, ...
		)
		panel.levelplot(
			x = x, y = y, z = z, subscripts = subscripts, region = FALSE,
			at = c(1,2,3), contour = TRUE,
			labels = list(cex = .75), label.style = 'align', ...
		)
	}
)
```

What about the individual $\mathrm{p}K_a$ values?

```{r, fig.height = 10}
pnw <- cbind(pna, t(mapply(function(est, true) {
	rowMeans((t(est) - true)^2)
}, pn$pK_est, pn$pK_true)))
levelplot(
	sqrt(pK1) + sqrt(pK2) + sqrt(pK3) ~ noise_ratio + n_points | noise_kind,
	aggregate(
		pnw[c('pK1','pK2','pK3')],
		pnw[c('noise_ratio','n_points','noise_kind')],
		mean
	),
	group = noise_kind, scales = list(log = 10),
	main = quote(
		'Standard deviation of '*pK[a]*' estimate'
	),
	xlab = 'Noise-to-signal ratio',
	ylab = 'Number of known points',
	cuts = 255, col.regions = hcl.colors(256, 'Terrain 2'),
	panel = function(x, y, z, subscripts, at, contour, region, ...) {
		panel.levelplot(
			x = x, y = y, z = z, subscripts = subscripts,
			at = at, contour = FALSE, region = TRUE, ...
		)
		panel.levelplot(
			x = x, y = y, z = z, subscripts = subscripts, region = FALSE,
			at = pretty(z), contour = TRUE,
			labels = list(cex = .75), label.style = 'align', ...
		)
	}
)
```

One more way to quantify the error. Let's take a look at every model
fitted so far, group by experiment conditions (noise-to-signal ratio,
number of points, noise distribution) and obtain the 95% highest
absolute deviation between estimated and true value of individual
$\mathrm{p}K_a$s. This will save us from having to recalculate standard
deviations of the estimates into confidence bands. Bonus points if the
error turs out to be independent from the noise distribution.

```{r, fig.height = 10}
nw <- cbind(n, t(abs(mapply(`-`, n$pK_est, n$pK_true))))
levelplot(
	pK1 + pK2 + pK3 ~ noise_ratio + n_points | noise_kind,
	aggregate(
		nw[c('pK1','pK2','pK3')],
		nw[c('noise_ratio','n_points','noise_kind')],
		quantile, .95
	),
	group = noise_kind, scales = list(log = 10),
	main = quote(
		{95^'th'}*' percentile of absolute '*pK[a]*' estimation error'
	),
	xlab = 'Noise-to-signal ratio',
	ylab = 'Number of known points',
	cuts = 255, col.regions = hcl.colors(256, 'Terrain 2'),
	panel = function(x, y, z, subscripts, at, contour, region, ...) {
		panel.levelplot(
			x = x, y = y, z = z, subscripts = subscripts,
			at = at, contour = FALSE, region = TRUE, ...
		)
		panel.levelplot(
			x = x, y = y, z = z, subscripts = subscripts, region = FALSE,
			at = pretty(z), contour = TRUE,
			labels = list(cex = .75), label.style = 'align', ...
		)
	}
)
```

Hard to say whether the noise distribution really makes a difference,
but $\mathrm{p}K_{a3}$ can be estimated reliably, $\mathrm{p}K_{a2}$ may
require a lot of points and good agreement between experiment and model,
and $\mathrm{p}K_{a1}$ is absolutely hopeless. But we knew about the latter.
